---
title: "Activity of Single Anonymous Person"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

### Loading Packages and Reading Data

The following code loads the data table package and the ggplot2 package which are required for the analysis.
```{r Reading Packages and data, results='asis',message=FALSE}
require(data.table)
require(ggplot2)
activity<-fread("activity.csv",key = "interval")
activity[,date:=as.Date(date,"%Y-%m-%d")]
knitr::kable(summary(activity))
```

### Total Number of Steps
Lets examine the total daily steps the subject takes each day.

```{r Total Steps}
totalSteps<-activity[,.(total.steps = sum(steps,na.rm = T)),by = date]
a<- ggplot(totalSteps)+
    theme(plot.title = element_text(hjust = 0.5))
a+geom_histogram(aes(total.steps),fill = 'red',colour = 'black',bins = 20)+
    ggtitle("Total Steps per Day")+
    xlab("Total Steps per Day")
```
  
As seen the most common number of steps per day is 0. This is due to a number of days having only missing values which will be addressed in a subsequent section. 

On average the subject takes `r round(totalSteps[,mean(total.steps,na.rm = T)],0)` steps per day and the on the median day takes `r totalSteps[,median(total.steps,na.rm = T)]` steps.

### Average Day

Next we will exam the subject's average day.  to do this we will create a graph of how meny steps the subject takes in each 5 minute interval.
```{r Daily Pattern}
 averageSteps<- activity[,.(average.steps = mean(steps,na.rm = T)), by = interval]
b<- ggplot(averageSteps)+
    theme(plot.title = element_text(hjust = 0.5))
b+geom_line(aes(interval,average.steps), colour = 'blue')+
    ggtitle("Average Steps Per Interval")+
    ylab("Average Steps")+
    xlab("Interval")
```

As seen in the Graph above the subject is sedentary in the mornings until around interval 500, and the subject takes them most average steps on interval `r averageSteps[average.steps == max(average.steps),interval]`.

### Missing Values

Now we will address the `r activity[,sum(is.na(steps))]` missing values by imputing the average number of steps taking during the same interval on all days were the value is not missing.
```{r Impute Missing Values}
setkey(averageSteps,interval)

activity[averageSteps, steps := ifelse(is.na(steps)
                                       ,as.integer(round(average.steps,0))
                                       ,steps)]

totalStepsm<-activity[,.(total.steps = sum(steps,na.rm = T)),by = date]
a<- ggplot(totalStepsm)+
    theme(plot.title = element_text(hjust = 0.5))
a+geom_histogram(aes(total.steps),fill = 'red',colour = 'black',bins = 20)+
    ggtitle("Total Steps per Day")+
    xlab("Total Steps per Day")
```

As expected the new distribution as shifted to the right since we added steps were there used to be missing values.  The new average is `r round(totalStepsm[,mean(total.steps,na.rm = T)],0)` which is `r round(totalStepsm[,mean(total.steps,na.rm = T)],0)-round(totalSteps[,mean(total.steps,na.rm = T)],0)`  higher then it was when we omited the missing values.  The new median is `r round(totalStepsm[,median(total.steps,na.rm = T)],0)` which is `r round(totalStepsm[,median(total.steps,na.rm = T)],0)-round(totalSteps[,median(total.steps,na.rm = T)],0)`  higher then it was when we omited the missing values. The without the missing values the distrubition is closer to a normal distribution.     

### Weekends vs Weekdays

Finally we will examine how the sujbect's weekdays differ from thier weekends.

```{r Weekends v. Weekdays}
activity[,Day :=as.factor(ifelse(grepl("^S",weekdays(date))
                           ,"Weekend"
                           ,"Weekday"))]

averageSteps<- activity[,.(average.steps = mean(steps,na.rm = T))
                        , by = list(interval,Day)]
b<- ggplot(averageSteps)+
    theme(plot.title = element_text(hjust = 0.5))
b+geom_line(aes(interval,average.steps), colour = 'blue')+
    facet_grid(Day~.)+
    xlab("Interval")+
    ylab("Average Number of Steps")+
    ggtitle("Average Steps on Weekends vs. Weekdays")
```
  
The subject takes longer to become active on weekends, however they take more steps in most intervals.
